---
title: "Baysesian Network on Casual Inference on femsmoke dataset"
output:
  html_document:
    df_print: paged
  html_notebook: default
  pdf_document: default
---


```{r}
library(faraway)
library(bnlearn)
library(dplyr)
```


directed acyclic graph (DAG) representation

femsmoke is a data frame with 28 observations on the following 4 variables. It records female smokers and non-smokers, their age group, and whether the subjects were dead or still alive after 20 years.

y- number of observed count for given combination

smoker-a factor with levels yes no

dead-a factor with levels yes no

age-a factor with agegroup levels 18-24 25-34 35-44 45-54 55-64 65-74 75+

Death is the targeted varibale. We are interested in finding the casual relationship between death and smoking and age.


```{r}
femsmoke
```


```{r}
femsmoke <- as_tibble(femsmoke) %>%
  print(n = Inf)
```

When we only study only on smoke~dead, we will find that non-smokers have a higher rate of death than smokers, which would be counter our intution and fact that smoking is bad for health.This is because smoke is confounded with age. Age also has a siginificant influence on death. Younger people smoke more than elder people and younger people are more likely to live when the data is gathered.


```{r}
ct <- xtabs(y ~ smoker + dead, data = femsmoke)
```


```{r}
ct1 <- xtabs(y ~ age + smoker, data = femsmoke)
```


```{r}
prop.table(ct1, 1)
```


```{r}
ct2 <- xtabs(y ~ smoker+ age, data = femsmoke)
```

```{r}
ct2
```


```{r}
prop.table(ct2, 2)
```


```{r}
prop.table(ct, 1)
```

Next,we will add age variable:

```{r}
xtabs(y ~age, data = femsmoke)
```

```{r}
prop.table(xtabs(y ~age, data = femsmoke))
```



```{r}
prop.table(xtabs(y ~ smoker + age, data = femsmoke), 2)
```





```{r}
prop.table(xtabs(y ~ dead+smoker, data = femsmoke), 2)
```



```{r}
library(tigerstats)
xtabs(y ~dead + smoker + age, data = femsmoke) %>%prop.table(c(3,2)) %>% round(2)
```







Bayesian Network

Knowledge:Age can influece whether you smoke or not but smoke cannot influence your age. age can influence death. Smoke can influence death. 


If we ignore age influence on death when design casual diagram:
```{r}
dag <- empty.graph(nodes = c("A","D","S"))
arc.set <- matrix(c("A","S",
                    "S","D"),
                  byrow = TRUE,ncol = 2,dimnames=list(NULL,c("from","to")))

arcs(dag) <- arc.set
A.lv <- c("18-24","25-34","35-44","45-54","55-64","65-74","75+")
S.lv <- c("yes","no")
D.lv <- c("yes","no")
A.prob <- array(c(0.09,0.21,0.18,0.16,0.18,0.13,0.05),dim=7,dimnames=list(A=A.lv))
S.prob <- array(c(0.47, 0.53, 0.44, 0.56, 0.47, 0.53, 0.63, 0.37, 0.49, 0.51, 0.22, 0.78, 0.17, 0.83),dim=c(2,7),dimnames = list(S=S.lv,A=A.lv))
D.prob <- array(c(0.24,0.76,0.31,0.69),dim=c(2,2),dimnames = list(D=D.lv,S=S.lv))

cpt=list(D=D.prob,S=S.prob,A=A.prob)

bn <-custom.fit(dag, cpt)
```

```{r}
graphviz.plot(dag)
```


```{r}
Smoked <- mutilated(bn, list(S = "yes"))
Non_Smoked <- mutilated(bn, list(S = "no"))
```

```{r}
Smoked_Death <- mean(rbn(Smoked,50000)["D"]=="yes")
Non_Smoked_Death <- mean(rbn(Non_Smoked,50000)["D"]=="yes")
Smoked_Death-Non_Smoked_Death
```

When using knowledge building diagram, if we ignore confounding effect, we can get the casual relationship wrong.




```{r}
alt.dag <- empty.graph(nodes = c("A","D","S"))
arc.set <- matrix(c( "A","S",
                     "A","D",
                    "S","D"),
                  byrow = TRUE,ncol = 2,dimnames=list(NULL,c("from","to")))

arcs(alt.dag) <- arc.set
```

```{r}
graphviz.plot(alt.dag)
```

```{r}
A.lv <- c("18-24","25-34","35-44","45-54","55-64","65-74","75+")
S.lv <- c("yes","no")
D.lv <- c("yes","no")
A.prob <- array(c(0.09,0.21,0.18,0.16,0.18,0.13,0.05),dim=7,dimnames=list(A=A.lv))
S.prob <- array(c(0.47, 0.53, 0.44, 0.56, 0.47, 0.53, 0.63, 0.37, 0.49, 0.51, 0.22, 0.78, 0.17, 0.83),dim=c(2,7),dimnames = list(S=S.lv,A=A.lv))
D.prob <- array(c(0.04,0.96,0.02,0.98,
                  0.02,0.98,0.03,0.97,
                  0.13,0.87,0.06,0.94,
                  0.21,0.79,0.15,0.85,
                  0.44,0.56,0.33,0.67,
                  0.81,0.19,0.78,0.22,
                  1,0,1,0),dim=c(2,2,7),dimnames = list(D=D.lv,S=S.lv,A=A.lv))
cpt=list(D=D.prob,A=A.prob,S=S.prob)
bn1<-custom.fit(alt.dag,cpt)
```

```{r}
Smoked <- mutilated(bn1, list(S = "yes"))
Non_Smoked <- mutilated(bn1, list(S = "no"))
Smoked_Death <- mean(rbn(Smoked,50000)["D"]=="yes")
Non_Smoked_Death <- mean(rbn(Non_Smoked,50000)["D"]=="yes")
Smoked_Death-Non_Smoked_Death
```


Smoke causes death increase 5%


```{r}
table(Cars93$Type, Cars93$Origin)
```


## transfer femsmoke to each record of a patient so that the data can be put into Agenarisk

```{r}
tab=xtabs(y ~dead + smoker + age, data = femsmoke)
fem=setNames(tidyr::uncount(as.data.frame(tab), Freq), c("dead", "smoker","age"))
```


```{r}
write.csv(fem,"~/Downloads/fem.csv",row.names=FALSE)
```

